---
title: "Reproducible Research: Peer Assessment 1"
author: Fernando DePaolis
output: 
  html_document:
    keep_md: true
---
```{r, echo=FALSE, message=F, warning=F}
library(plyr)
library(dplyr)
library(Hmisc)
library(ggplot2)
```

## Loading and preprocessing the data
```{r, results='asis'}
activity<- read.csv ("activity.csv", header=TRUE, sep=",")
attach(activity)
```


## What is mean total number of steps taken per day?
```{r}
summary(activity)
##Mean total number of steps per day
steps.day<-aggregate(steps~date, FUN=sum)
mean(steps.day$steps)
median(steps.day$steps)
```

###Histogram of Steps taken per day
```{r}
hist(steps.day$steps, breaks = 20, 
     main = paste("Steps per day"), xlab = "Steps", 
     density = 30, angle=45, 
     border="black", col="lightblue")
```

## What is the average daily activity pattern?

```{r}
steps.interval <- aggregate(steps~interval, FUN=mean)
mean(steps.interval$steps)
```


###Line chart of Steps Taken
```{r}
ggplot(mapping = aes(x = unique(activity$interval), y = steps.interval$steps)) + 
        geom_line(col="indian red") + scale_x_continuous("5min intervals (similar to military time)", breaks = seq(min(unique(activity$interval)), max(unique(activity$interval)), 100)) + 
        scale_y_continuous("Mean Steps per Period") + ggtitle("Average Number of Steps Taken by Interval") + theme(axis.text.x = element_text(size = rel(.75), angle = 00))
```


## Imputing missing values
[based on solution at http://grokbase.com/t/r/r-help/0969yezzan/r-how-to-substitute-missing-values-nas-by-the-group-means]
```{r, results='hide',warning=FALSE, message=FALSE}

activity2 <- activity

as.numeric(apply(activity2, 1, function (x)
        ifelse( is.na(x[2]),
        tapply(activity2$steps,activity2$interval,
        mean, na.rm=TRUE)[x[1]],
        x[2])))

attach(activity2)
steps.day2<-aggregate(steps~date, FUN=sum)
```
```{r, results='show',warning=FALSE, message=TRUE}
mean(steps.day2$steps)
median(steps.day2$steps)
```
###Histogram "AFTER" imputing data to missing valus
```{r}
hist(steps.day2$steps, breaks = 20, 
     main = paste("Steps per day"), xlab = "Steps", 
     density = 30, angle=45, 
     border="black", col="lightblue")
```

## Are there differences in activity patterns between weekdays and weekends?
###Assigning days of the week (weekday/weekend) to dates
```{r}
activity2$day.type <- ifelse(weekdays(as.Date(activity2$date)) %in% c("Saturday", "Sunday"),"weekend","weekday")
```

###Line chart panels by type of day of the week
```{r}
activity3 <- activity2 %>% group_by(interval, day.type) %>% summarise(meanSteps = mean(steps, na.rm = TRUE))

ggplot(data = activity3, mapping = aes(x = interval, y = meanSteps)) + 
        geom_line(col="indian red") + facet_grid(day.type ~ .) + scale_x_continuous("5min intervals (similar to military time)", breaks = seq(min(activity3$interval), max(activity3$interval), 100)) + 
        scale_y_continuous("Mean Steps per Period") + ggtitle("Average Number of Steps Taken by Interval") + theme(axis.text.x = element_text(size = rel(.75), angle = 00))
```

